---
title: "GRANAR-MECHA USERGUIDE"
authors: Adrien Heymans, Marco D'Agostino, Valentin Couvreur, Guillaume Lobet
format: html
output-file: 'index.html'
editor: source
bibliography: bibliography.bib
---

```{r preambule, include=F}
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # Set working directory to the current file
rm(list = ls())

libs <- c("pander", "readxl", "reticulate", "xml2", "tidyverse", "dplyr", "devtools", "viridis", "ggplot2", "gridExtra", "sf", "deldir", "packcircles")
# "riot"
#==========================================================================
for(pk in libs) {
  if (!requireNamespace(pk, quietly = TRUE)) {
    # If not installed, install the package
    install.packages(pk)
  }
  
  library(pk, character.only = T)
}

# Load utilitarian functions
source("src/functions.R")
source("src/update_params.R")
source("src/update_MECHA_inputs.R")
source("src/MECHA_inputs.R")
```

# What is GRANAR?

GRANAR stands for Generator of Root ANAtomy in R. The GRANAR model is able to generate complete cell networks of root cross-sections using a small set of root anatomical parameters These root anatomical features can typically be gathered using software image analysis such as [ImageJ](https://imagej.net/Welcome) [@Schneider]. Once the root anatomy is created, the cell network can be saved as an eXtended Markup Language (XML) file. The structure of the XML files created by GRANAR are identical to the ones created by [CellSeT](https://www.nottingham.ac.uk/research/groups/cvl/software/cellset.aspx) [@Pound] (a software which digitize root anatomical network).

# How to use GRANAR?

## Load functions

GRANAR is available as an R package. The *'granar'* package can be found on [GitHub](https://github.com/granar/granar). You can clone the repository with Github desktop and load all functions :

```{r GRANAR, include  = T, eval = T}
# Load GRANAR with local functions
granar_path <- "../granar/R/"
for(i in list.files(granar_path)){
  # print(i)
  source(paste0(granar_path, i))
}
```

Alternatively, GRANAR can be installed localy by simply executing the following line in the [R](https://cran.r-project.org/) environment:

```{r Github, eval = FALSE}
# install_github("granar/granar") 
# library(granar)
```

## Run GRANAR

To run GRANAR, we use :

-   The function *read_param_xml()* loads parameters from xml files
-   The function *create_anatomy()* generates the virtual cross section with input params
-   The function *plot_anatomy()* displays the generated anatomy

### Monocot (wheat)

```{r run GRANAR, results=F, message=FALSE}
param1 <- read_param_xml(path = "Params/Wheat_F_Atkinson_2017.xml")
sim1   <- suppressWarnings(create_anatomy(parameters = param1))
plot_anatomy(sim1)
```

### Maize with aerenchyma

```{r}
param <- read_param_xml(path = "Params/Zea_mays_Heymans_2019.xml")
sim   <- suppressWarnings(create_anatomy(parameters = param))
plot_anatomy(sim)
```

### Rice

```{r}
param <- read_param_xml(path = "Params/Rice_D_Atkinson_2017.xml")
sim   <- suppressWarnings(create_anatomy(parameters = param))
plot_anatomy(sim)
```

### Arabodopsis

```{r}
param <- read_param_xml(path = "Params/Arabidopsis.xml")
sim <- suppressWarnings(create_anatomy(parameters = param))
plot_anatomy(sim)
```

### Primary growth dicot (tomato)

```{r run GRANAR2, results=F, message=FALSE}
param2 <- read_param_xml(path = "src/Params/Tomato_primary.xml")
sim2 <- suppressWarnings(create_anatomy(parameters = param2))
plot_anatomy(sim2)
```

### Secondary growth dicot (tomato)

```{r run GRANAR3, results=F, message=FALSE}
param3 <- read_param_xml(path = "src/Params/Tomato_secondary.xml")
sim3 <- suppressWarnings(create_anatomy(parameters = param3))
plot_anatomy(sim3)
```

To save virtual anatomies, we can use the function *write_anatomy_xml()* :

```{r write_anatomy, eval = F}
write_anatomy_xml(sim = sim1, path = "Anatomies/Wheat.xml")
write_anatomy_xml(sim = sim2, path = "Anatomies/Tomato_primary.xml")
write_anatomy_xml(sim = sim3, path = "Anatomies/Tomato_secondary.xml")
```

```{r load anatomy, include = F, eval = F}
sim1 <- get_root_section(path = "Anatomies/Wheat.xml")

ggplot(sim1) +
    # geom_polygon(aes_string("x", "y", group="id_cell", fill=col), colour="white") +
    geom_polygon(aes(x = x, y = y, fill = type, group = id_cell),
                 color = "white") +
    theme_classic() +
    coord_fixed() +
    theme(axis.line=element_blank(),
          axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())
```

# MECHA

MECHA is the solver of local conductivities (Couvreur 2018)

Set up steps to run it in local :

1.  Open **ANACONDA NAVIGATOR**

2.  Create a new environment with **python 3.7** (in this work it is called **GRANAR-MECHA**)\
    `conda create -n GRANAR-MECHA python=3.7`

3.  Load that environment

    `conda activate GRANAR-MECHA`

4.  Install MECHA requirements in the environement :\
    `conda install numpy`\
    `conda install scipy`\
    `conda install matplotlib`\
    `conda install lxml`\
    `conda install pip`\
    `pip install networkx==1.9.1`

By default, MECHA loads the cross section named *current_root.xml* in the *cellsetdata* folder, so we need to copy the anatomy we created to that folder

## Load and setup MECHA params

```{r}
# Update parameters
MECHA_path   <- "src/MECHA/"                              # Path to MECHA file directory
project_path <- "Projects/granar/"                          # Path to MECHA projects
input_path   <- paste0("in/")                 # Path where MECHA will take parameters  
output_path  <- paste0("out/")                       # Path where MECHA will store outputs
read_path    <- paste0(MECHA_path, project_path, input_path) # Path where MECHA_input_list will be created from XML. Set template if different.

# LOAD XML AS A BIG LIST

# Read XML input files as templates
MECHA_inputs_list <- read_MECHA_inputs(
  read_path, 
  names = list(
    General    = "General.xml",
    Hydraulics = "Hydraulics.xml",
    Hormones   = "Hormones_Carriers.xml",
    BC         = "BC.xml",
    Geometry   = "Geometry.xml")
  ) 
```

## Modify MECHA params by hand

MECHA has a lot of parameters, most of them have a standard value coming from literature.

You can modify at will the parameters , like such :

```{r}
# MODIFY PARAMETERS BY HAND
MECHA_inputs_list$Hydraulics$param$kAQPrange$kAQP$value <- 0.001 # Default value : 1e-4 cm/hPa/d 

# WRITE XML's
write_MECHA_inputs(input_list = MECHA_inputs_list, 
                   path = paste0(MECHA_path, project_path, input_path),
                   names = list(
                     General    = "General.xml",
                     Hydraulics = "Hydraulics.xml",
                     Hormones   = "Hormones_Carriers.xml",
                     BC         = "BC.xml",
                     Geometry   = "Geometry.xml")
                   )
```

## Set up anatomy to send to MECHA

```{r, eval = F}
anatomy_file <- "Tomato_primary.xml"

file.copy(from = paste0("Anatomies/", anatomy_file),
          to = "src/MECHA/cellsetdata/", overwrite = T)

file.rename(paste0("src/MECHA/cellsetdata/", anatomy_file), 
            to = "src/MECHA/cellsetdata/current_root.xml")

sim1 <- get_root_section(path = paste0("Anatomies/", anatomy_file))
ggplot(sim1) +
    # geom_polygon(aes_string("x", "y", group="id_cell", fill=col), colour="white") +
    geom_polygon(aes(x = x, y = y, fill = type, group = id_cell),
                 color = "white") +
    theme_classic() +
    coord_fixed() +
    theme(axis.line=element_blank(),
          axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())
```

## Run MECHA

By default, MECHA can simulate different scenarios : \*

-   **B0 : \[Special case for protophloem\] No apoplastic barriers**

-   **B1 : Endodermal Casparian strip (radial walls)**

-   **B2 : Endodermal suberization except at passage cells**

-   **B3 : Endodermis full suberization**

-   **B4 : Endodermis full suberization and exodermal Casparian strip (radial walls)**

-   **B5 : Endo- and exodermal Casparian strips (radial walls)**

-   **B6 : Exodermis full suberization and endodermal Casparian strip (radial walls)**

-   **B7 : Exodermal Casparian strip (radial walls)**

-   **B8 : Endo and exodermal suberization**

-   **B9 : Lignin Cap and endodermal CS**

Here we choose to compute scenarios 1, 3 and 8. You can change the simulated scenarios in *MECHA/Projects/granar/in/Maize_Geometry.xml* in line 19.

Let's run MECHA :

```{r MECHA, eval = F}
use_condaenv("GRANAR-MECHA")
sys <- import("sys", convert=T)

unlink("src/MECHA/Projects/granar/out/Tomato/Root/Project_Test/results/", force = T)
try(py_run_file("src/MECHA/MECHA_example.py"))
```

We can visualise the ratio of symplastiv vs. apoplastic flows for each scenario : nb ; Graphs are broken with Secondary Growth scenario

```{r, results=F, message=FALSE, warning=FALSE}
rp <- "src/MECHA/Projects/granar/out/Tomato/Root/Project_Test/results/"

Flux1 <- ApoSymp(paste0(rp,"Macro_prop_1,0.txt")) 
Flux2 <- ApoSymp(paste0(rp,"Macro_prop_2,1.txt")) 
Flux3 <- ApoSymp(paste0(rp,"Macro_prop_3,2.txt")) 


plot_water_flux(Flux1, apobar = 1)
plot_water_flux(Flux2, apobar = 2)
plot_water_flux(Flux3, apobar = 3)
```

We can load radial conductivity $k_r$ and axial conductance $K_x$ :

```{r, message=FALSE}
# Load Barriers names and IDs
Barrier_types <- read_xlsx("src/MECHA/Barrier_types.xlsx") %>% 
  mutate(ID = as.factor(ID)) %>% 
  mutate(Barrier_name = paste0("En.", Endodermis, " - Ex.", Exodermis))

# Use custom function to load outputs of MECHA
conds <- load_MECHA_out(path = "src/MECHA/Projects/granar/out/Tomato/Root/Project_Test/results/")

# Process and select the outputs
conds <- conds %>% 
  mutate(ID = as.factor(ID),
         kr = as.numeric(kr),
         Kx = as.numeric(Kx)) %>% 
  left_join(Barrier_types, by = "ID") %>% 
  filter(Endodermis %in% c("-", "CS", "Sub") & Exodermis %in% c("-", "CS", "Sub"))

conds <- conds[order(conds$kr),]
# pander(conds)

# Quick and uggly Plot
ggplot(data = conds) +
  geom_col(aes(x = Barrier_name, 
               y = kr,
               fill = Endodermis,
               color = Exodermis
               ),
           size = 2,
           width = 0.85
           ) +
  scale_color_viridis(discrete = T) +
  scale_fill_viridis(discrete = T) +
  theme_test() +
  # facet_wrap(~Exodermis) +
  theme(axis.text.x = element_text(angle = 90))
```
